%!TEX root = ../document.tex

\section[Conclusion (Author: Patrick Schilf)]{Conclusion}
This document presents our final prototype of a Eclipse plugin that supports software developers in their workflow. Using SAP's HANA our tool allows programmers to instantly investigate the data that they write code for as well as focus on their tasks from one integrated environment. Our prototype integrates well in the Eclipse environment, giving the developer information about runtime, relevant and real Data Contexts as well as potential upcoming issues at a glimpse. The programmer benefits from these information as they are presented earlier and quicker than in a traditional development cycle. As a major challenge regarding a technical implementation of our prototype we see the filtering and clustering to import particular relevant Data Contexts into the environment. However, we see this feature as an essential part of our solution, as it enables programmers to not only work more time efficient but also to ensure the quality of their code. Once working, the feature enables the developer to even interact with the on demand result sets, e.g. by altering result relations or switching between contexts, to investigate the influence on the program flow. In the final chapter of this document we give insights of what considerations to make when implementing this feature.

Our prototype reflects the needs of our Persona Amber, that we developed throughout the seminar. The understanding and observing phases of the design thinking process helped us to quickly dig into the problem domain and define these needs. In the further phases of ideation, prototyping, and testing we were able to integrate the repeatedly gained feedback and new insights in a final solution concept.